CALVIS: Chest, wAist and peLVIS circumference from 3D human body meshes as ground truth for deep learning

### 2019-08-02 PaperCVFAD2019/egpaper_for_review.tex
mesh signature extrema

Intuitively, we would like to measure the chest circumference bellow the arms 
at the widest part of the torso and the waist circumference at the 
narrowest part after the chest but above the hips. Similarly, the hip 
circumference is measured often around the widest part of hips and buttocks. We 
can formalize this intuition by considering the cross-sectional length of the 
2D intersection curves (\cite{book.compu.topo}) along the y-axis. Moreover, we 
can 
use a plane $\boldsymbol{\pi}$ parallel to the floor to intersect the mesh. 
Since $\mathcal{M}$ is 
triangulated, the boundary of this 
intersection is a collection of segments $s_i$. Therefore, we can 
determine the boundary length as at point j
\begin{align}
\mathcal{BL}_j = \sum_{i = 1}^{i = n}s_i
\end{align}

Note that for cross sections where $\boldsymbol{\pi}$ intersects the legs two 
disconnected curves will appear. That is by no means a problem because they are 
still a collection of segment with non-zero length.

\begin{figure}[t]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Figure_1.eps}
	\end{center}
	\caption{Human Body Mesh Signature for male and female meshes}
	\label{fig:hbm_signature}
\end{figure}

Next, starting from the top of the bounding box we cut the mesh into evenly 
spaced slices $m$. For every slice j, we compute $BL_j$. Finally, we can 
construct the human body \textbf{mesh signature} $\mathcal{MS}$ of mesh 
$\mathcal{M}$ in the zero pose $\vec{\theta}^*$

\begin{align}
\mathcal{MS}(\mathcal{M}, \vec{\theta}^*, j) = \mathcal{BL}_j, \, j \in \{1, 
\cdots, m\}
\end{align}

Figure \ref{fig:hbm_signature} shows the $\mathcal{MS}$ for two meshes. The 
function resembles a rotated silhouette of the human body and exhibits a number 
of critical points.

\subsection{Critical Points on Mesh Signature}

>>>>

Intuitively, we would like to measure the chest circumference bellow the arms 
at the widest part of the torso and the waist circumference at the 
narrowest part after the chest but above the hips. Similarly, the hip 
circumference is measured often around the widest part of hips and buttocks. We 
can formalize this intuition by considering the cross-sectional length of the 
2D intersection curves (\cite{book.compu.topo}) along the y-axis. Moreover, we 
can 
use a plane $\boldsymbol{\pi}$ parallel to the floor to intersect the mesh. 
Since $\mathcal{M}$ is 
triangulated, the boundary of this 
intersection is a collection of segments $s_i$. Therefore, we can 
determine the boundary length by specifying a point along the y-axis $j$ as
\begin{align}
\mathcal{BL}(\mathcal{M}, \boldsymbol{\pi}, j) = \sum_{i = 
1}^{i = n}s_i
\end{align}

Note that for cross sections where $\boldsymbol{\pi}$ intersects the legs, two 
disconnected curves will appear. That is by no means a problem because they are 
still a collection of segment with non-zero length.

\begin{figure}[t]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Figure_1.eps}
	\end{center}
	\caption{Human Body Mesh Signature for male and female meshes. The 
		function resembles a rotated silhouette of the human body and exhibits 
		several \textit{extrema}.}
	\label{fig:hbm_signature}
\end{figure}

Next, we assemble the mesh slide length vector $\vec{\mathcal{L}}$. Starting 
from the 
top $j_t$ of the bounding box we cut the mesh into 
evenly 
$m$-meters spaced slices until we reach the bounding box bottom $j_b$. For 
every 
slice at point $j$, we compute $\mathcal{BL}(j)$.

\begin{align}
\vec{\mathcal{L}} = \mathcal{BL}(j), \, j 
\in 
\{j_t, j_t+m, \cdots, j_b\}
\end{align}

Finally, we can construct the human body \textbf{mesh signature} $\mathcal{MS}: 
\mathbb{Z}^+ \to \mathbb{R}$ of mesh $\mathcal{M}$ sliced iterative by plane 
$\boldsymbol{\pi}$ every $m$-meters by setting its value to the value of 
$\vec{\mathcal{L}}$ at 
slide index $x$

\begin{align}\label{eq:mesh_signature}
\mathcal{MS}(\mathcal{M}, \boldsymbol{\pi}, m) = \vec{\mathcal{L}}(x), \, x \in 
\{0, 1, \cdots\}
\end{align}

Figure \ref{fig:hbm_signature} shows the $\mathcal{MS}$ of two meshes (male 
and female) for 
$m=0,001$ and plane $\boldsymbol{\pi}$ parallel to the floor (with normal $(0, 
1,0)$) using the library trimesh \cite{trimesh}. The 
function resembles a rotated silhouette of the human body and exhibits several 
\textit{extrema}.

\subsection{Mesh Signature Extrema}

By comparing neighboring values of the mesh signature $\mathcal{MS}$ as defined 
in equation \ref{eq:mesh_signature} we find several \textit{extrema}. In 
general, we use these \textit{extrema} as features to 
calculate the human dimensions. More specifically, we assume that:
\begin{enumerate}
	\item The global maximum $M_g$ represents the length of the largest path 
	around the arms.
	\item The local maximum $M_{hc}$ with largest value other than $M_g$ 
	immediately posterior to it represents the hip circumference.
	\item The local minimum $M_{wc}$ posterior to $M_g$ but prior to $M_{hc}$ 
	represents hip circumference.
	\item The local maximum $M_{cc}$ posterior to $M_g$ but prior to $M_{wc}$ 
	represents chest circumference.
\end{enumerate}
